\documentclass{article}

\usepackage{ipprocess}
% Use longtable if you want big tables to split over multiple pages.
% \usepackage{longtable}
\usepackage[utf8]{inputenc} 
% Babel package is used to translate keywors to a specific language. 
% The option "brazil" defines portuguese-brazil as default language.
% Babel is also useful for hiphenation.
\usepackage[brazil]{babel}
\usepackage{mathtools}
% \usepackage[LGRgreek]{mathastext}

\sloppy

\title{MUSA Core Processor}

\graphicspath{{./pictures/}} 
\makeindex
\begin{document}
\DocumentTitle{Especificação do Projeto}
\Project{Processador LAPI DOpaCA LAMBA}
\Organization{Universidade Estadual de Feira de Santana}
\Version{Build 2.1}
 \capa

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Historico de Revisoes
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{\center Histórico de Revisões}
  \vspace*{1cm}
  \begin{table}[ht]
    \centering
    \begin{tabular}[pos]{|m{2cm} | m{7.2cm} | m{3.8cm}|} 
      \hline
      \cellcolor[gray]{0.9}
      \textbf{Date} & \cellcolor[gray]{0.9}\textbf{Descrição} & \cellcolor[gray]{0.9}\textbf{Autor(s)}\\ \hline
      \hline
      \small xx/xx/xxxx & \small <Descrição> & \small <Autor(es)> \\ \hline
      \small 08/12/2015 & \small Teste & \small Filipe Boaventura \\ \hline
      \small xx/xx/xxxx &
      \begin{small}
        \begin{itemize}
          \item Exemplo de;
          \item Revisões em lista;
        \end{itemize}
      \end{small} & \small <Autor(es)> \\ \hline 
    \end{tabular}
  \end{table}

\newpage
  
\tableofcontents
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Concent
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \section{Introducão}
  \subsection{Propósito}
  O objetivo desse documento é definir as especificações para o processador \textbf{LAPI DOpaCA LAMBA}, incluindo o seu montador e simulador. Aqui são definidos os parametros de implementação do processador que incluem: modos de operação, arquitetura do conjunto de instruções (ISA), e a descrição dos registradoes e flags.
   Este documento está organizado da seguinte forma: Visão Geral na seção 2, Arquitetura do Conjunto de Instruções seção 3, Montador seção 4, Simulador seção 5.
  
  \subsection{Document Outline Description}
  Este documento está organizado da seguinte forma:
	
	\begin{itemize}
	  \item Section \color{black}{\ref{sec:design_overview}}: Apresenta uma visão geral do processador. 
	  \item Section \color{black}{\ref{sec:isa}}: Especifica a arquitetura do conjunto de instruções.
	  \item Section \color{black}{\ref{sec:register}}: Describes the general purpose registers.
	\end{itemize}
		
  \subsection{Acronyms and Abbreviations}
  
  \FloatBarrier
  \begin{table}[H]
    \begin{center}
      \begin{tabular}[pos]{|m{2cm} | m{9cm}|} 
	\hline
	\cellcolor[gray]{0.9}\textbf{Acronym} & \cellcolor[gray]{0.9}\textbf{Description} \\ \hline
	    RISC   & Reduced Instruction Set Computer \\ \hline
	    GPR    & General Purpose Registers \\ \hline
	    FPGA   & Field Gate Programmable Array \\ \hline
	    GPPU   & General Purpose Processing Unit \\ \hline
	    SPRAM  & Single-port RAM \\ \hline	    
	    HDL    & Hardware Description Language \\ \hline
	    CPU    & Central Processing Unit \\ \hline
	    ISA    & Instruction Set Architecture \\ \hline
	    ALU    & Arithmetic and Logic Unit \\ \hline
	    PC     & Program Counter \\ \hline
	    RF     & Flags Register \\ \hline
	    CST    & Constant Value \\ \hline
      \end{tabular}
    \end{center}
  \end{table}
  
  \newpage
  \section{Visão Geral}
  \label{sec:design_overview}
  O LAPI DOpaCA LAMBA é um processador RISC que apresenta um conjunto de 42 instruções organizadas em X grupos funcionais e possiu três modos de endereçamento. O processador possui 16 registradores de propósito geral em seu banco de registradores. O LAPI DOpaCA LAMBA suporta operações lógicas e aritiméticas básicas, excerto multiplicação e divisão. 
  
 \subsection{Perspectives}
    
  The MUSA 32-bit core release targets modern FPGA devices with embedded SPRAM memory, and is intended to be deployed as a GPPU soft IP-core. Its architecture is designed under MIPS original implementation, presenting a slightly reduced ISA.
  
  \subsection{Características Principais}
  
  \begin{itemize}
   \item Arquitetura de 32 bits;
   \item 16 registradores de propósito geral;
   \item Arquitetura de conjunto de instruções (ISA) RISC;
   \item ISA composta por 42 instruções;
   \item Organização de dados Big-Endian;
   \item Três modos de endereçamento: immediato, por registrador, e base-deslocamento;
   
  \end{itemize}

  \subsection{Non-functional Requirements}
  
  \begin{itemize}
   \item The FPGA prototype should run in a Terasic ALTERA Cyclone III (EP3C25F324) Development platform;
   \item The design must be described using Verilog-HDL;
   \item The tests must be writen in both Verilog-HDL or SystemVerilog;
   \item A set of test programs must be provided in order to validate the implementation;
  \end{itemize}

  
  \newpage
  \section{Conjunto de Instruções}
  \label{sec:isa}
  O Processador LAPI DOpaCA LAMBA possui 42 instruções de 32 bits de largura que estão organizadas nos seguintes grupos funcionais:
  \begin{itemize}
  \item ALU;
  \item Constantes;
  \item Transferências de controle (jump);
  \item Memória (load e store);
  \item NOPs;
  \item Condição de término.
  \end{itemize}
  
  In MUSE load/store architecture, operations are performed through operands held into the register file (GPR Bank) and main memory is accessed only through load and store instructions. Signed and unsigned integers of 16-bits are supported by loads that either sign-extend or zero-extend the data loaded into the register.  
  
  \subsection{CPU Load and Store Instructions}
  
 \FloatBarrier
  \begin{table}[H]
    \begin{center}
      \label{tab:load_store_instructions}
      \begin{tabular}[pos]{| l | l | l | l |} \hline 	
      \multicolumn{1}{|c|}{\cellcolor[gray]{0.9}\textbf{Mnemonic}} &
      \multicolumn{1}{c|}{\cellcolor[gray]{0.9}\textbf{Operands}} &  
      \multicolumn{1}{c|}{\cellcolor[gray]{0.9}\textbf{Realization}} &
      \multicolumn{1}{c|}{\cellcolor[gray]{0.9}\textbf{Description}} \\ \hline
	LW  & $RD,RS1,I_{16}$ & $RD$ \textleftarrow $MEM[RS1+I_{16}]$ & Load word from data memory.	\\ \hline
	SW  & $RS1,RS2,I_{16}$ & $MEM[RS1+I_{16}]$ \textleftarrow $RS2$ & Store word into data memory. 		\\ \hline
      \end{tabular}
    \end{center}
  \end{table}  
  
  
  \subsection{Computational Instructions}
  
  The MUSA provides 32-bit arithmetic operations. The computational instructions uses ALU two-operand instructions. These are signed versions of the following operations:

  \begin{itemize}
   \item Add
   \item Subtract
   \item Multiply
   \item Divide
  \end{itemize}
  
 \FloatBarrier
  \begin{table}[H]
    \begin{center}
      \begin{tabular}[pos]{| l | l | l | m{5cm} |} \hline 	
      \multicolumn{1}{|c|}{\cellcolor[gray]{0.9}\textbf{Mnemonic}} &
      \multicolumn{1}{c|}{\cellcolor[gray]{0.9}\textbf{Operands}} &
      \multicolumn{1}{c|}{\cellcolor[gray]{0.9}\textbf{Realization}} &
      \multicolumn{1}{m{5cm}|}{\cellcolor[gray]{0.9}\textbf{Description}} \\ \hline
	ADD  	& $RD$, $RS1, RS2$ 	& $RD$ \textleftarrow $RS1 + RS2$ 	& Add two words. 			\\ \hline
	SUB 	& $RD$, $RS1, RS2$ 	& $RD$ \textleftarrow $RS1 - RS2$ 	& Subtract two words. 		\\ \hline
	MUL 	& $RD$, $RS1, RS2$ 	& $RD$ \textleftarrow $RS1 * RS2$ 	& Multiply two words.		\\ \hline
	DIV   	& $RD$, $RS1, RS2$ 	& $RD$ \textleftarrow $RS1 / RS2$ 	& Divide two words.		\\ \hline
	AND  	& $RD$, $RS1, RS2$ 	& $RD$ \textleftarrow $RS1 \odot RS2$ 	& Logical AND.		\\ \hline
	OR 	    & $RD$, $RS1$ 	    & $RD$ \textleftarrow $RS1 \oplus RS2$ 	& Logical OR.	\\ \hline
	ADDI  	& $RD$, $RS1, I_{16}$ 	& $RD$ \textleftarrow $RS1 + I_{16}$ 	& Add a stored word and an \mbox{immediate} value. \\ \hline
	SUBI 	& $RD$, $RS1, I_{16}$ 	& $RD$ \textleftarrow $RS1 - I_{16}$ 	& Subtract a stored word and an immediate value.\\ \hline
	ANDI 	& $RD$, $RS1, I_{16}$ 	& $RD$ \textleftarrow $RS1 \odot I_{16}$ 	& Logical AND of a stored word and an immediate value		\\ \hline
	ORI   	& $RD$, $RS1, I_{16}$ 	& $RD$ \textleftarrow $RS1 \oplus I_{16}$ 	& Logical OR of a stored word and an immediate value.		\\ \hline	
	CMP 	& $RD$, $RS1$ 	    & -- 					            & Compares $RD$ and $RS1$ and set the flags.\\ \hline
	NOT 	& $RD$ 	            & $RD$ \textleftarrow $\sim RD$     & Logical NOT.	\\ \hline
      \end{tabular}
    \end{center}
  \end{table} 
  
  \subsection{Jump/Branch Instructions}

  \FloatBarrier
    \begin{table}[H]
      \begin{center}
		\begin{tabular}[pos]{| l | l | l | l |} \hline 	
	\multicolumn{1}{|c|}{\cellcolor[gray]{0.9}\textbf{Mnemonic}} &
	\multicolumn{1}{c|}{\cellcolor[gray]{0.9}\textbf{Operands}} &
	\multicolumn{1}{c|}{\cellcolor[gray]{0.9}\textbf{Characteristic}} &
	\multicolumn{1}{c|}{\cellcolor[gray]{0.9}\textbf{Description}} 			\\ \hline
	  JR  	& $R$ 		& Unconditional & Jump to destination. 			\\ \hline
	  JPC 	& $I_{28}$ 	& Unconditional & Jump to destination PC-relative. 	\\ \hline
	  BRFL 	& $RF$, $CST$ 	& Conditional 	& Jump to destination if $RF == CST$.\\ \hline
	  CALL   	& $R$		& Unconditional & Subroutine call.	\\ \hline
	  RET  	& -- 		& Unconditional & Subroutine return 			\\ \hline
	\end{tabular}
      \end{center}
    \end{table} 
  
  \subsection{No Operation Instruction}
  The \textit{No Operation} instruction (\texttt{NOP}) is used to control the instruction flow or to insert delays (stalls) into the datapath, such as when computing the result of a jump/branch instruction. When using a NOP instruction after a branch/jump instruction it is so named a \textbf{branch delay slot}.
  
  \subsection{End of Opperations}
  The HALT instruction (system stop) must be implemented as a \texttt{L: j L} (a unconditional branch to the current address)
  
  \section{Internal General Purpose Registers}
  \label{sec:register}
  The current MUSA architecture provides 32 fixed-point general purpose registers of 32 bits: $R_0$ to $R_{31}$. The $R_0$ have special use for the hardware always returning zero, no matter what software attempts to store to it.

\end{document}
